### 2022-02-17 plots/carto_indee_covid.Rmd
Add cartography

---
title: "Carto"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

```{r cars}
library(sf)
library(tidyverse)
library(plotly)
library(viridis)
library(leaflet)
library(ggspatial)
library(gridExtra)
library(htmltools)
```

```{r}
poly = st_read("georef-france-epci-millesime.shp")
poly_com = st_read("georef-france-commune-arrondissement-municipal-millesime.shp")

REV_brut = read_csv("REV_brut.csv")
LOG_brut = read.csv("LOG_brut.csv")

```

```{r}
poly$epci_code <- as.numeric(levels(poly$epci_code))[poly$epci_code]
poly_com$com_arm_cur <- as.numeric(levels(poly_com$com_arm_cur))[poly_com$com_arm_cur]

plot_REV = poly %>%
  inner_join(REV_brut, by = c("epci_code" = "codgeo")) %>%
  rename(codgeo = epci_code) %>%
  plyr::rbind.fill(poly_com %>%
                     inner_join(REV_brut, by = c("com_arm_cur" = "codgeo")) %>%
                     rename(codgeo = com_arm_cur))

plot_LOG = poly %>%
  inner_join(LOG_brut, by = c("epci_code" = "codgeo")) %>%
  rename(codgeo = epci_code) %>%
  plyr::rbind.fill(poly_com %>%
                     inner_join(LOG_brut, by = c("com_arm_cur" = "codgeo")) %>%
                     rename(codgeo = com_arm_cur))
```

# Represent vaccination rate

```{r}
vacct = read_csv("vacc_rates4.csv")
```

```{r}
plot_vacc = poly %>%
  rename(codgeo = epci_code) %>%
  plyr::rbind.fill(poly_com %>%
                     rename(codgeo = com_arm_cur)) %>%
  inner_join(vacct)
```

```{r}
carto_vacc = function(wk){
  vacc_sem = plot_vacc %>% filter(week == wk)

  min_r = min(vacc_sem$taux_cumu_1_inj, na.rm = T)
  max_r = max(vacc_sem$taux_cumu_1_inj, na.rm = T)

  set_scale = function(x, min_r, max_r){
    return((x-min_r)/(max_r-min_r))
  }

  min_p = set_scale(min(vacc_sem[vacc_sem$epci_name == "Métropole du Grand Paris",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)
  max_p = set_scale(max(vacc_sem[vacc_sem$epci_name == "Métropole du Grand Paris",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)

  min_l = set_scale(min(vacc_sem[vacc_sem$epci_name == "Métropole de Lyon",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)
  max_l = set_scale(max(vacc_sem[vacc_sem$epci_name == "Métropole de Lyon",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)

  min_m = set_scale(min(vacc_sem[vacc_sem$epci_name == "Métropole d'Aix-Marseille-Provence",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)
  max_m = set_scale(max(vacc_sem[vacc_sem$epci_name == "Métropole d'Aix-Marseille-Provence",]$taux_cumu_1_inj, na.rm = T), min_r, max_r)

  p1 = vacc_sem %>%
    ggplot() +
    aes(fill = taux_cumu_1_inj) +
    scale_fill_viridis(name = paste("Vaccination Rate\nWeek", wk),option = "cividis") +
    geom_sf(size = 0)  +
    theme_void() +
    theme(legend.position = "top") +
    coord_sf(crs = st_crs(4326)) +
    labs(title = "France")

  p2 = vacc_sem %>%
    filter(epci_name == "Métropole du Grand Paris") %>%
    ggplot() +
    aes(fill = taux_cumu_1_inj) +
    scale_fill_viridis(option = "cividis", begin = min_p, end = max_p) +
    geom_sf(size = 0)  +
    theme_void() +
    theme(legend.position = "None") +
    coord_sf(crs = st_crs(4326)) +
    labs(title = "Paris")

  p3 = vacc_sem %>%
    filter(epci_name == "Métropole de Lyon") %>%
    ggplot() +
    aes(fill = taux_cumu_1_inj) +
    scale_fill_viridis(option = "cividis", begin = min_l, end = max_l) +
    geom_sf(size = 0)  +
    theme_void() +
    theme(legend.position = "None") +
    coord_sf(crs = st_crs(4326)) +
    labs(title = "Lyon")

  p4 = vacc_sem %>%
    filter(epci_name == "Métropole d'Aix-Marseille-Provence") %>%
    ggplot() +
    aes(fill = taux_cumu_1_inj) +
    scale_fill_viridis(option = "cividis", begin = min_m, end = max_m) +
    geom_sf(size = 0)  +
    theme_void() +
    theme(legend.position = "None") +
    coord_sf(crs = st_crs(4326)) +
    labs(title = "Marseille-Aix")

  return(grid.arrange(p1, p2, p3, p4,
                      widths = c(4,1,1,1),
                      layout_matrix = rbind(c(1, 2),
                                            c(1, 3),
                                            c(1, 4))))
}
```

```{r}
carto_vacc(27)
carto_vacc(31)
carto_vacc(35)
carto_vacc(52)
```

# Represent Unemployment

```{r}
min_r = min(plot_REV$Unemployment_Benef, na.rm = T)
max_r = max(plot_REV$Unemployment_Benef, na.rm = T)

set_scale = function(x, min_r, max_r){
  return((x-min_r)/(max_r-min_r))
}

min_p = set_scale(min(plot_REV[plot_REV$epci_name == "Métropole du Grand Paris",]$Unemployment_Benef, na.rm = T), min_r, max_r)
max_p = set_scale(max(plot_REV[plot_REV$epci_name == "Métropole du Grand Paris",]$Unemployment_Benef, na.rm = T), min_r, max_r)

min_l = set_scale(min(plot_REV[plot_REV$epci_name == "Métropole de Lyon",]$Unemployment_Benef, na.rm = T), min_r, max_r)
max_l = set_scale(max(plot_REV[plot_REV$epci_name == "Métropole de Lyon",]$Unemployment_Benef, na.rm = T), min_r, max_r)

min_m = set_scale(min(plot_REV[plot_REV$epci_name == "Métropole d'Aix-Marseille-Provence",]$Unemployment_Benef, na.rm = T), min_r, max_r)
max_m = set_scale(max(plot_REV[plot_REV$epci_name == "Métropole d'Aix-Marseille-Provence",]$Unemployment_Benef, na.rm = T), min_r, max_r)

p1 = plot_REV %>%
  ggplot() +
  aes(fill = Unemployment_Benef) +
  scale_fill_viridis(name = "Unemployment\nBenefits",option = "magma") +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "top") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "France")

p2 = plot_REV %>%
  filter(epci_name == "Métropole du Grand Paris") %>%
  ggplot() +
  aes(fill = Unemployment_Benef) +
  scale_fill_viridis(option = "magma", begin = min_p, end = max_p) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Paris")

p3 = plot_REV %>%
  filter(epci_name == "Métropole de Lyon") %>%
  ggplot() +
  aes(fill = Unemployment_Benef) +
  scale_fill_viridis(option = "magma", begin = min_l, end = max_l) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Lyon")

p4 = plot_REV %>%
  filter(epci_name == "Métropole d'Aix-Marseille-Provence") %>%
  ggplot() +
  aes(fill = Unemployment_Benef) +
  scale_fill_viridis(option = "magma", begin = min_m, end = max_m) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Marseille-Aix")

grid.arrange(p1, p2, p3, p4,
             widths = c(4,1,1,1),
             layout_matrix = rbind(c(1, 2),
                                   c(1, 3),
                                   c(1, 4)))
```

# Reprensent Overcrowding Rate

```{r}
min_r = min(plot_LOG$Overcrowding_rate, na.rm = T)
max_r = max(plot_LOG$Overcrowding_rate, na.rm = T)

set_scale = function(x, min_r, max_r){
  return((x-min_r)/(max_r-min_r))
}

min_p = set_scale(min(plot_LOG[plot_LOG$epci_name == "Métropole du Grand Paris",]$Overcrowding_rate, na.rm = T), min_r, max_r)
max_p = set_scale(max(plot_LOG[plot_LOG$epci_name == "Métropole du Grand Paris",]$Overcrowding_rate, na.rm = T), min_r, max_r)

min_l = set_scale(min(plot_LOG[plot_LOG$epci_name == "Métropole de Lyon",]$Overcrowding_rate, na.rm = T), min_r, max_r)
max_l = set_scale(max(plot_LOG[plot_LOG$epci_name == "Métropole de Lyon",]$Overcrowding_rate, na.rm = T), min_r, max_r)

min_m = set_scale(min(plot_LOG[plot_LOG$epci_name == "Métropole d'Aix-Marseille-Provence",]$Overcrowding_rate, na.rm = T), min_r, max_r)
max_m = set_scale(max(plot_LOG[plot_LOG$epci_name == "Métropole d'Aix-Marseille-Provence",]$Overcrowding_rate, na.rm = T), min_r, max_r)

p1 = plot_LOG %>%
  ggplot() +
  aes(fill = Overcrowding_rate) +
  scale_fill_viridis(name = "Overcrowding\nRate",option = "plasma") +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "top") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "France")

p2 = plot_LOG %>%
  filter(epci_name == "Métropole du Grand Paris") %>%
  ggplot() +
  aes(fill = Overcrowding_rate) +
  scale_fill_viridis(option = "plasma", begin = min_p, end = max_p) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Paris")

p3 = plot_LOG %>%
  filter(epci_name == "Métropole de Lyon") %>%
  ggplot() +
  aes(fill = Overcrowding_rate) +
  scale_fill_viridis(option = "plasma", begin = min_l, end = max_l) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Lyon")

p4 = plot_LOG %>%
  filter(epci_name == "Métropole d'Aix-Marseille-Provence") %>%
  ggplot() +
  aes(fill = Overcrowding_rate) +
  scale_fill_viridis(option = "plasma", begin = min_m, end = max_m) +
  geom_sf(size = 0)  +
  theme_void() +
  theme(legend.position = "None") +
  coord_sf(crs = st_crs(4326)) +
  labs(title = "Marseille-Aix")

grid.arrange(p1, p2, p3, p4,
             widths = c(4,1,1,1),
             layout_matrix = rbind(c(1, 2),
                                   c(1, 3),
                                   c(1, 4)))
```

# Interactive maps

## Unemployment Benefits

```{r}
# Remove missing data
plot_REV = plot_REV %>% filter(!is.na(Unemployment_Benef))

# set colours
pal = scales::gradient_n_pal(colours = c("lightblue", "black", "orange"),
                             values = c(min(plot_REV$Unemployment_Benef),
                                        median(plot_REV$Unemployment_Benef, na.rm = T),
                                        max(plot_REV$Unemployment_Benef)))

# we are representing the two layers (EPCI & Municipalities) separately
plot_REV_epci = poly %>%
  inner_join(REV_brut, by = c("epci_code" = "codgeo")) %>%
  rename(codgeo = epci_code)
plot_REV_com = poly_com %>%
  inner_join(REV_brut, by = c("com_arm_cur" = "codgeo")) %>%
  rename(codgeo = com_arm_cur)

# apply colours
plot_REV_epci$hex = pal(plot_REV_epci$Unemployment_Benef)
plot_REV_com$hex = pal(plot_REV_com$Unemployment_Benef)

# Build label
plot_REV_epci = plot_REV_epci %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", epci_name, "</p><p>Unemployment Benefits: ", round(Unemployment_Benef, 2), "%</p>")) %>%
  ungroup()
plot_REV_com = plot_REV_com %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", com_arm_nam, "</p><p>Unemployment Benefits: ", round(Unemployment_Benef, 2), "%</p>")) %>%
  ungroup()

leaflet() %>%
  addTiles() %>%
  addPolygons(data = plot_REV_epci,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_REV_epci$lab, HTML)
              label = ~paste0("Unemployment Benefits: ", round(Unemployment_Benef, 2), "%")
              ) %>%
  addPolygons(data = plot_REV_com,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_REV_com$lab, HTML)
              label = ~paste0("Unemployment Benefits: ", round(Unemployment_Benef, 2), "%")
              ) %>%
  addProviderTiles("Stamen.TonerLite")
```

## Overcrowding rate

```{r}
plot_LOG = plot_LOG %>% filter(!is.na(Overcrowding_rate))
pal = scales::gradient_n_pal(colours = c("green", "black", "red"),
                             values = c(min(plot_LOG$Overcrowding_rate),
                                        median(plot_LOG$Overcrowding_rate, na.rm = T),
                                        max(plot_LOG$Overcrowding_rate)))

plot_LOG_epci = poly %>%
  inner_join(LOG_brut, by = c("epci_code" = "codgeo")) %>%
  rename(codgeo = epci_code)
plot_LOG_com = poly_com %>%
  inner_join(LOG_brut, by = c("com_arm_cur" = "codgeo")) %>%
  rename(codgeo = com_arm_cur)

plot_LOG_epci$hex = pal(plot_LOG_epci$Overcrowding_rate)
plot_LOG_epci = plot_LOG_epci %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", epci_name, "</p><p>Overcrowding rate: ", round(Overcrowding_rate, 2), "%</p>")) %>%
  ungroup()
plot_LOG_com$hex = pal(plot_LOG_com$Overcrowding_rate)
plot_LOG_com = plot_LOG_com %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", com_arm_nam, "</p><p>Overcrowding rate: ", round(Overcrowding_rate, 2), "%</p>")) %>%
  ungroup()

leaflet() %>%
  addTiles() %>%
  addPolygons(data = plot_LOG_epci,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_LOG_epci$lab, HTML)
              label = ~paste0("Overcrowding rate: ", round(Overcrowding_rate, 2), "%")
              ) %>%
  addPolygons(data = plot_LOG_com,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_LOG_com$lab, HTML)
              label = ~paste0("Overcrowding rate: ", round(Overcrowding_rate, 2), "%")
              ) %>%
  addProviderTiles("Stamen.TonerLite")
```

## Vaccination rate

```{r}
plot52 = plot_vacc %>%
  filter(!is.na(taux_cumu_1_inj) & week == 52)

pal = scales::gradient_n_pal(colours = c("purple", "black", "yellow"),
                             values = c(min(plot52$taux_cumu_1_inj),
                                        median(plot52$taux_cumu_1_inj, na.rm = T),
                                        max(plot52$taux_cumu_1_inj)))

plot_vacc_epci = poly %>%
  inner_join(vacct %>% filter(week == 52), by = c("epci_code" = "codgeo")) %>%
  rename(codgeo = epci_code)
plot_vacc_com = poly_com %>%
  inner_join(vacct %>% filter(week == 52), by = c("com_arm_cur" = "codgeo")) %>%
  rename(codgeo = com_arm_cur)

plot_vacc_epci$hex = pal(plot_vacc_epci$taux_cumu_1_inj)
plot_vacc_com$hex = pal(plot_vacc_com$taux_cumu_1_inj)

plot_vacc_epci = plot_vacc_epci %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", epci_name, "</p><p>Vaccination rate week 52: ", round(taux_cumu_1_inj*100, 2), "%</p>")) %>%
  ungroup()
plot_vacc_com = plot_vacc_com %>%
  group_by(codgeo) %>%
  mutate(lab = paste0("<p>", com_arm_nam, "</p><p>Vaccination rate week 52: ", round(taux_cumu_1_inj*100, 2), "%</p>")) %>%
  ungroup()

leaflet() %>%
  addTiles() %>%
  addPolygons(data = plot_vacc_epci,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_vacc_epci$lab, HTML)
              label = ~paste0("Vaccination rate week 52: ", round(taux_cumu_1_inj*100, 2), "%")
              ) %>%
  addPolygons(data = plot_vacc_com,
              color = ~hex,
              weight = 1,
              opacity = 0.4,
              fillOpacity = 0.5,
              stroke = T,
              highlight = highlightOptions(opacity = 0.8,
                                           fillOpacity = 0.9,
                                           bringToFront = T),
              # label = lapply(plot_vacc_com$lab, HTML)
              label = ~paste0("Vaccination rate week 52: ", round(taux_cumu_1_inj*100, 2), "%")
              ) %>%
  addProviderTiles("Stamen.TonerLite")
```