### 2022-02-22 code/florence.Rmd
first commit

---
title: "VaccinationIndicators"
author: "FD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initializations

```{r}
rm(list = ls())

# Package for odds ratios
library("epitools")
# Package for colors
library("RColorBrewer")
# Package for dynamical plots
library("plotly")

#library("igraph")
```

# Load data

## Indicators

```{r}
# The data have been dealt with in "0_INSEE_predictors.R" and saved as RData
load("../data/predictors.RData")

# Load region information for the different EPCI
regions <- read.csv("../data/geographic/EPCI_composition-communale.csv")
head(regions)
# Create dictionary 
dic.reg <- regions$REG
names(dic.reg) <- regions$EPCI
```

```{r}

```

## Vaccination data

Note: The files can be downloaded again (for updates) by running "0_loadAmeliData.R"

### EPCI

```{r}
vaccEPCI <- read.csv("../data/ameli_vaccination/vaccEPCI.csv", sep = ";")
head(vaccEPCI)

# Add region information
vaccEPCI$region <- dic.reg[as.character(vaccEPCI$epci)]
table(vaccEPCI$region, useNA = "ifany")

# Note: NAs correspond to unattributed locations
unique(vaccEPCI[is.na(vaccEPCI$region), "epci"])

# Exclude DROM using regional code
vaccEPCI <- vaccEPCI[vaccEPCI$region > 9, ]

# Exclude Metropoles for which we have Communes detail
epci.PLM <- c(200054781, # Paris
              200046977, # Lyon
              200054807  # Marseille
              )
vaccEPCI <- vaccEPCI[which(!is.element(vaccEPCI$epci, epci.PLM)), ]

# Rename some fields for future merging
names(vaccEPCI)[c(3, 4)] <- c("codgeo", "libelle_geo")
```

### Communes

```{r}
vaccCom <- read.csv("../data/ameli_vaccination/vaccCom.csv", sep = ";")
head(vaccCom)
# Rename some fields for future merging
names(vaccCom)[c(3, 4)] <- c("codgeo", "libelle_geo")

# Merge vaccination data into a single dataset
vacc <- rbind(vaccEPCI[, 1:16], vaccCom)

# Clean memory
rm(vaccCom, vaccEPCI)

# Make sure that population sizes are numeric
vacc$population_carto <- as.numeric(vacc$population_carto)

```

## Dates

```{r}
# Get all dates in the vaccination dataset
vaccDates <- sort(unique(vacc$date))

# Define specific dates
date1 <- "2021-07-11" # Just before pass sanitaire announcement
date2 <- "2021-08-08" # Just before pass sanitaire comes into force
date3 <- "2021-08-29" # Before back to school
date4 <- "2022-01-02" # Last date of the year
```

# Analysis

## Some definitions of common functions

### Quantiles

Conversion to binary values depending on a quantile

```{r}
# Convert to binary values depending on a chosen threshold
quantileConvert <- function(v, q){
  # v vector to be converted
  # q chosen quantile (between 0 and 1)

  # Compute the threshold value
  # Rem: need to remove NAs
  qval <- quantile(x = v, probs = q, na.rm = TRUE)

  # Return TRUE/FALSE depending on the position wrt the value, and change into 1/0
  out <- ifelse(v < qval, 0, 1)

  out
}

# Convert to binary values, with respect to 2 threshold values
quantilesConvert <- function(v, qmin = 0.1, qmax = 0.9){
  # v vector to be converted
  # qmin lower threshold value, data below it are coded 0
  # qmax higher threshold value, data above it are coded 1
  # and data between qmin and qmax are coded NA

  # Compute the threshold value
  # Rem: need to remove NAs
  qval <- quantile(x = v, probs = c(qmin, qmax), na.rm = TRUE)

  out <- rep(NA, length(v)) # Initialize output vector
  out[v < qval[1]] <- 0 
  out[v > qval[2]] <- 1

  out
}
```

### Treatment of the indicators

-  Discretization of the data with respect to quantiles

```{r}
# Function to discretize wrt to quantiles
discretizeQ <- function(v, prbs = seq(0, 1, by = 0.05)){
  # v     vector to discretize
  # prbs  positions of the boundaries

  # Define quantiles
  qq <- c(quantile(v, probs = prbs, na.rm = TRUE))

  # Extend the extreme values to make sure that they are included later
  qq[1] <- 0.5 * qq[1] # 0%
  qq[length(qq)] <- 1.5 * qq[length(qq)] # 100%

  # Add noise to avoid issues when data are too close
  noise <- c(0, runif(length(prbs)-2, max = 10^(-6)), 0)
  qq <- sort(qq + noise)

  # Put the values of v in the different classes
  out <- cut(v, breaks = qq, right = FALSE, labels = FALSE)
# table(out, useNA = "ifany")
  out
}
```

```{r, eval = FALSE}
# Test
for(pred in unique(names(dat.all)[-1])){
  vv <- dat.all[, pred]
  plot(discretizeQ(vv, prbs = seq(0, 1, by = 0.1)), vv, main = pred)
}
```

### Treatment of the vaccination data

-  Adjust the data

```{r}
adjustVacc <- function(r1, varVacc){
  # r1  vaccination dataset for a given date, with all ages, but not "TOUT_AGE"
  # varVacc  the vaccination variable that we want to investigate ("1_inj" or "termine")

  # Compute distribution of the population by age class
  #   Population size of each age classs
  distPop <- aggregate(r1$population_carto, by = list(classe_age = r1$classe_age), FUN = sum, na.rm = TRUE)
  #   Proportions
  distPop$propAgeClass <- distPop$x / sum(distPop$x)

  # Merge with vaccination data
  rr1 <- merge(r1, distPop, by = "classe_age")

  # Compute adjusted rates, by weighting each by each by the proportion in the overall population
  s1 <- aggregate(data.frame(taux_cumu = rr1[, paste0("taux_cumu_", varVacc)] * rr1$propAgeClass, population_carto = rr1$population_carto), by = list(codgeo = rr1$codgeo), FUN = sum, na.rm = TRUE)
  # Compute adjusted numbers of vaccinated people
  s1$n_cumu <- round(s1$taux_cumu * s1$population_carto)

  # Return the dataframe
  s1
}
```

-  Remove children

```{r}
adultVacc <- function(r1, varVacc){
  # r1  vaccination dataset for a given date, with all ages, but not "TOUT_AGE"
  # varVacc  the vaccination variable that we want to investigate ("1_inj" or "termine")

  # Remove children age class
  rr1 <- r1[which(r1$classe_age != "00-19"), ]

  # Compute sums of effectifs by codgeo
  s1 <- aggregate(data.frame(n_cumu = rr1[, paste0("effectif_cumu_", varVacc)], population_carto = rr1$population_carto), by = list(codgeo = rr1$codgeo), FUN = sum, na.rm = TRUE)

  # Compute taux
  s1$taux_cumu <- s1$n_cumu / s1$population_carto

  # Return the dataframe
  s1
}
```

## Other functions

```{r}
invlogit <- function(x) exp(x)/(1 + exp(x))

```

## Parameters

```{r}
parms <- expand.grid(clAge = sort(unique(vacc$classe_age)), 
                     varVacc = c("1_inj", "termine"), 
                     varPred = names(dat.all[, -1]), 
                     thedate = c(date1, date2, date3, date4), stringsAsFactors = FALSE)
```

## Odds ratios

### Function definitions

```{r computeOR}
# Function to compute odds ratio and p value
computeOR <- function(pred, result, method = "fisher"){
  # pred   vector of the predictor, dichotomized
  # result vector of the result, dichotomized
  # method one of c("midp", "fisher", "wald", "small"), see ?oddsratio

  # Compute contingency table
  contTable <- table(pred, result)

  # Compute OR and pval
  or <- oddsratio(x = contTable, method = method)
  # Format the output
  out <- c(N = or$data[3, 3], or$measure[2, ], or$p.value[2, ])

  #print(or) # For debugging and double-checking
  out
}
```

```{r}
getOR <- function(clAge, varVacc, varPred, thedate, thrVacc = 0.25, thrPred = 0.5){
  # clAge age class
  # varVacc variable for the vaccination data
  # varPred variable for the predictor
  # thedate date of the vaccination data
  # thrVacc threshold value for dichotomizing vaccination data
  # thrPred threshold value for dichotomizing predictor data

  # Subset vaccination data
  r1 <- vacc[vacc$classe_age == clAge & vacc$date == thedate, ]

  # Merge with predictor data
  m1 <- merge(r1, dat.all[, c("codgeo", varPred)], by = "codgeo")

  # Dichotomize the vectors
  result <- quantileConvert(m1[, paste0("taux_cumu_", varVacc)], thrVacc)
  pred <- quantileConvert(m1[, varPred], thrPred)

  # Compute OR
  computeOR(pred, result)
#  output <- c(date = thedate, varVacc = varVacc, clAge = clAge, varPred = varPred, computeOR(pred, result))
#  output
}
```

```{r computeORCheck}
# For debugging
# clAge <- "TOUT_AGE"
# varVacc <- "taux_cumu_1_inj"
# varPred <- "Overcrowding_rate"
# thedate <- date1
# dataTransform <- "discretize"
```

### Computation

```{r}
# Compute for all combinations that we want
# (Takes a few minutes)
nrow(parms)
output <- matrix(0, ncol = 7, nrow = nrow(parms))
for(i in 1:nrow(parms)){
  cat(i, "")
  output[i, ] <- do.call(getOR, parms[i, ])
}

# For debugging
getOR("TOUT_AGE", "1_inj", "Overcrowding_rate", date1, thrVacc = 0.25, thrPred = 0.5)

# Add the parameter information
mm <- cbind(parms, output)

names(mm) <- c("clAge", "varVacc", "varPred", "thedate", "N", "estimate", "lower", "upper", "midp.exact", "fisher.exact", "chi.square")

# Save the output
save(mm, file = "output_OR.RData")
```

## Anova

### Function definitions

```{r}
anovaVacc <- function(clAge, varVacc, varPred, thedate, thrPred = c(0.1, 0.9)){
  # clAge age class
  # varVacc variable for the vaccination data
  # varPred variable for the predictor
  # thedate date of the vaccination data
  # thrPred threshold value(s) for dichotomizing predictor data
  #         if one value, all data are binarized as 0: v<q, 1: v>=q
  #         if two values, 0: v<q1, NA: q1<= v <=q2, 1: v>q2

  # Subset vaccination data
  r1 <- vacc[vacc$classe_age == clAge & vacc$date == thedate, ]

  # Merge with predictor data
  m1 <- merge(r1, dat.all[, c("codgeo", varPred)], by = "codgeo")

  # Dichotomize the vector
  if(length(thrPred) == 1){
    m1$pred <- quantileConvert(m1[, varPred], thrPred)
  }else{
    m1$pred <- quantilesConvert(m1[, varPred], thrPred[1], thrPred[2])
  }
  m1$res <- m1[, paste0("taux_cumu_", varVacc)]

  aov(res ~ pred, data = m1)
}
```

### Computation

```{r}
mdl <- anovaVacc("TOUT_AGE", "1_inj", "Overcrowding_rate", date3)
(mdl$coefficients)[["pred"]]
summary(mdl)

effect3 <- rep(0, ncol(dat.all)-1)
for(i in 2:ncol(dat.all)){
  mdl <- anovaVacc("TOUT_AGE", "1_inj", names(dat.all)[i], date3)
  effect3[i-1] <- (mdl$coefficients)[["pred"]]
}

summary(mdl)
```

## Logistic regression

### Function definitions
```{r}
# Function to do a logistic regression on one of the variables
getLogReg <- function(varVacc, varPred, thedate, predTransform, vaccAge, by.prbs = 0.1, permutation = FALSE){
  # 
  # varVacc  takes values "1_inj" or "termine"
  # varPred  is the predictor value to be used
  # thedate  date at which the calculations are done
  # predTransform: how to deal with the predictor values 
  #                "standardize_msd", "standardize_max" or "discretize"
  # vaccAge: "raw_rates" if raw vaccination rates, without taking age into account
  #          "by_age" if model takes age into account
  #          "adjusted_rates" if adjusted vaccination rates
  #          "adults" all ages except 00-19
  # by.prbs  if "discretize", distance between breaks of the discretization

  # Check data entry
  stopifnot(is.element(predTransform, c("standardize_msd", "standardize_max", "discretize")))
  stopifnot(is.element(varVacc, c("1_inj", "termine")))
  stopifnot(is.element(vaccAge, c("raw_rates", "by_age", "adjusted_rates", "adults")))

  if(vaccAge == "raw_rates"){
    # Subset vaccination data; keep only ALL_AGE
    r1 <- vacc[vacc$classe_age == "TOUT_AGE" & vacc$date == thedate, ]
    # Rename columns
    r1$taux_cumu <- r1[, paste0("taux_cumu_", varVacc)]
    r1$n_cumu <- r1[, paste0("effectif_cumu_", varVacc)]
  }
  if(vaccAge == "by_age"){
    # Subset vaccination data; keep all age classes except ALL_AGE
    r1 <- vacc[vacc$classe_age != "TOUT_AGE" & vacc$date == thedate, ]
    # Rename columns
    r1$taux_cumu <- r1[, paste0("taux_cumu_", varVacc)]
    r1$n_cumu <- r1[, paste0("effectif_cumu_", varVacc)]
  }
  if(vaccAge == "adjusted_rates"){
    # Subset vaccination data; keep all age classes except ALL_AGE
    tmp <- vacc[vacc$classe_age != "TOUT_AGE" & vacc$date == thedate, ]
    r1 <- adjustVacc(tmp, varVacc)
  }
  if(vaccAge == "adults"){
    # Subset vaccination data; keep all age classes except ALL_AGE
    tmp <- vacc[vacc$classe_age != "TOUT_AGE" & vacc$date == thedate, ]
    r1 <- adultVacc(tmp, varVacc)
  }

  # Merge with predictor data
  m1 <- merge(r1, dat.all[, c("codgeo", varPred)], by = "codgeo")

  # Treat the predictor
  pp <- m1[, varPred]
  if(predTransform == "standardize_msd"){
    m1$pred.std <- (pp - mean(pp, na.rm = TRUE)) / sd(pp, na.rm = TRUE)
  }
  if(predTransform == "standardize_max"){
    pMax <- max(pp, na.rm = TRUE)
    pMin <- min(pp, na.rm = TRUE)
    m1$pred.std <- (pp - pMin) / (pMax - pMin)
  }
  if(predTransform == "discretize"){
    m1$pred.std <- discretizeQ(v = pp, prbs = seq(0, 1, by = by.prbs))
  }

  # Permutation if we want to (to test signif)
  if(permutation){
    permi <- sample(seq_len(nrow(m1)))
    m1$pred.std <- m1[permi, "pred.std"]
  }

  # Logistic model
  if(vaccAge == "by_age"){
    # Logistic model, with age effect
    mdl <- glm(cbind(n_cumu, population_carto - n_cumu) ~ pred.std + as.factor(classe_age) + pred.std * as.factor(classe_age), family = binomial(link = "logit"), data = m1)
  }
  if(is.element(vaccAge, c("raw_rates", "adjusted_rates", "adults"))){
    # Logistic model, no age effect (but potentially adjusted)
    mdl <- glm(cbind(n_cumu, population_carto - n_cumu) ~ pred.std, family = binomial(link = "logit"), data = m1)
  }

  # Extract values for pred.std
#  out <- summary(mdl)$coefficients["pred.std", ]
  # Return model
  mdl
}
```

### Computation, brouillon

```{r, eval = FALSE}
tmp <- getLogReg("1_inj", "Overcrowding_rate", thedate = date1, predTransform = "discretize", vaccAge = "raw_rates")
summary(tmp)
summary(getLogReg("1_inj", "Overcrowding_rate", thedate = date1, predTransform = "discretize", vaccAge = "raw_rates", permutation = TRUE))

tmp <- getLogReg("1_inj", "Overcrowding_rate", thedate = date1, predTransform = "standardize_msd", vaccAge = "raw_rates")
summary(tmp)

parmsLR <- expand.grid(varVacc = c("1_inj"), 
                     varPred = names(dat.nocorr[, -1]), 
                     thedate = c(date1, date2, date3, date4), 
#                     predTransform = c("standardize_msd", "standardize_max", "discretize"),
                     predTransform = c("discretize"),
                     vaccAge = c("raw_rates", "adjusted_rates", "adults"), 
                     permutation = c(FALSE, TRUE), 
                     stringsAsFactors = FALSE)
sort(names(dat.nocorr[, -1]))
nrow(parmsLR)
library("parallel")

outLR <- mclapply(seq_len(nrow(parmsLR)), function(i) do.call(getLogReg, parmsLR[i, ]))#,  mc.cores = 3)

# It takes a while with a list, but then we can check how fast it goes
outLR <- list()
for (i in seq_len(nrow(parmsLR))){
    cat(i, " ")
  outLR[[i]] <- do.call(getLogReg, parmsLR[i, ])
}

# Add line number in parms dataframe
parmsLR$i <- seq_len(nrow(parmsLR))

# Extract slope values
slopes <- unlist(lapply(seq_len(nrow(parmsLR)), function(i) summary(outLR[[i]])$coefficients["pred.std", "Estimate"]))

# Extract p values associated to the slopes
pvals <- unlist(lapply(seq_len(nrow(parmsLR)), function(i) summary(outLR[[i]])$coefficients["pred.std", "Pr(>|z|)"]))

# Compute decile values
tx01 <- unlist(lapply(seq_len(nrow(parmsLR)), function(i){cf <- outLR[[i]]$coefficients
invlogit(cf[1] + 1 * cf[2])}))

tx10 <- unlist(lapply(seq_len(nrow(parmsLR)), function(i){cf <- outLR[[i]]$coefficients
invlogit(cf[1] + 10 * cf[2])}))

# Put these values in a data frame
outputLR <- parmsLR
outputLR$slopes <- slopes
outputLR$pvals <- pvals
outputLR$tx01 <- tx01
outputLR$tx10 <- tx10

head(outputLR)

hist(-log(outputLR$pvals))

hist(outputLR$slopes)

hist(abs(outputLR$slopes))

head(outputLR[order(abs(outputLR$slopes), decreasing = TRUE), ])

for(predTransform in unique(outputLR$predTransform)){
  print("-----------------------------------------------------------------------------------")
  for(varVacc in unique(outputLR$varVacc)){
    for(thedate in unique(outputLR$thedate)){
      for(vaccAge in unique(outputLR$vaccAge)){
        for(permutation in c(TRUE, FALSE)){
          sub <- outputLR[outputLR$vaccAge == vaccAge & outputLR$thedate == thedate & outputLR$predTransform == predTransform & outputLR$varVacc == varVacc & outputLR$permutation == permutation, ]
          print(head(sub[order(abs(sub$slopes), decreasing = TRUE), ], n = 10))
          cat("\n")
          #        print(head(sub[order(abs(sub$tx10 - sub$tx01), decreasing = TRUE), ], n = 10))
          #        cat("\n")
          plot(abs(sub$slopes), ylim = c(0, max(c(0.07, max(abs(sub$slopes))))), 
               main = paste(predTransform, thedate, vaccAge, permutation))
        }
      }
    }
  }
}

vv <- dat.all$Unemployment_Benef
plot(discretizeQ(vv, seq(0, 1, by = 0.1)), vv)

invlogit <- function(x) exp(x)/(1 + exp(x))

outputLRD <- outputLR[which(outputLR$predTransform == "discretize"), ]

max(outputLRD$slopes)

save(outputLR, file = "outputLR_2022-02-19.RData")

```

Rationale for choosing how to treat the predictors:   
-  `standardize_msd` keeps extreme values, so still hard to compare distributions,
-  `standardize_max` also keeps extreme values, and if there are outliers, the rest of the distribution is completely condensed,  
so we will keep `discretize`.

```{r}
extractValuesOut <- function(outList, parmsMat){
  # Check dimensions
  stopifnot(length(outList) == nrow(parmsMat))

  # Initialize output
  m <- parmsMat

  # Add line number in parms dataframe
  m$i <- seq_len(nrow(parmsMat))

  # Extract slope values
  slopes <- unlist(lapply(seq_len(nrow(parmsMat)), function(i) summary(outList[[i]])$coefficients["pred.std", "Estimate"]))

# Extract p values associated to the slopes
  pvals <- unlist(lapply(seq_len(nrow(parmsMat)), function(i) summary(outList[[i]])$coefficients["pred.std", "Pr(>|z|)"]))

# Compute decile values
  tx01 <- unlist(lapply(seq_len(nrow(parmsMat)), function(i){cf <- outList[[i]]$coefficients
invlogit(cf[1] + 1 * cf[2])}))

  tx10 <- unlist(lapply(seq_len(nrow(parmsMat)), function(i){cf <- outList[[i]]$coefficients
invlogit(cf[1] + 10 * cf[2])}))

  # Compute SE
  preds <- predict(outList[[i]], newdata = data.frame(pred.std = 1:10), type = "response", se.fit = TRUE)
  se01 <- preds$se.fit[1]
  se10 <- preds$se.fit[10]

  # Put these values in a data frame
  m$slopes <- slopes
  m$pvals <- pvals
  m$tx01 <- tx01
  m$tx10 <- tx10
  m$se01 <- se01
  m$se10 <- se10
  m$difftx101 <- tx10 - tx01
  m
}
```

### Permutations

```{r permutationsSlopes, eval = FALSE}
# Combinations of parameters to test
parmsPerm <- expand.grid(varVacc = c("1_inj"), 
                     varPred = "French_nlty", 
                     thedate = c(date1, date2, date3, date4), 
                     predTransform = c("discretize"),
                     vaccAge = c("raw_rates", "adults"), 
                     permutation = c(TRUE), 
                     stringsAsFactors = FALSE)

# Number of replicates of each combination
nrep <- 1000

# Rewrite as table with replicates
#   List of the parameters, one element for one replicate
listParmsPerm <- list()
for(rep in 1:nrep) listParmsPerm[[rep]] <- parmsPerm
#   rbind them
parmsPermRep <- do.call(rbind, listParmsPerm)
#   clean memory
rm(listParmsPerm)
#   Check dimension
dim(parmsPermRep)

nperm <- nrow(parmsPerm) * nrep

# Compute model for all combinations and replicates
# It takes a while with a list, but then we can check how fast it goes
outPerm <- list()
for(rep in seq_len(nrep)){ # Replicates
  for (i in seq_len(nrow(parmsPerm))){ # Parameter values
    j <- (rep - 1) * nrow(parmsPerm) + i
    cat(j, " ")
    outPerm[[j]] <- do.call(getLogReg, parmsPerm[i, ])
  }
}

# Extract the key values
pPerm <- extractValuesOut(outList = outPerm, parmsMat = parmsPermRep)

save(pPerm, file = "randomPermutations_2022-02-21.RData")

# Clean memory 
rm(outPerm)
```

```{r}
length(outPerm)
dim(pPerm)  
head(pPerm)

hist(pPerm$slopes)
hist(abs(pPerm$slopes))

# Key values
quantile(abs(pPerm$slopes), 0.95)
quantile(abs(pPerm$slopes), 0.99)
quantile(abs(pPerm$slopes), 0.999)
max(abs(pPerm$slopes))

```

### Computation

```{r}
# Values to be tested
parmsLR <- expand.grid(varVacc = c("1_inj"), 
                     varPred = names(dat.nocorr[, -1]), 
                     thedate = c(date1, date2, date3, date4), 
                     predTransform = c("discretize"),
                     vaccAge = c("raw_rates", "adults"), 
                     permutation = c(FALSE), 
                     stringsAsFactors = FALSE)
```

```{r, eval = FALSE}
#outLR <- mclapply(seq_len(nrow(parmsLR)), function(i) do.call(getLogReg, parmsLR[i, ]))#,  mc.cores = 3)

nrow(parmsLR)

# It takes a while with a list, but then we can check how fast it goes
outLR <- list()
for (i in seq_len(nrow(parmsLR))){
    cat(i, " ")
  outLR[[i]] <- do.call(getLogReg, parmsLR[i, ])
}

# Extract the key values
pLR <- extractValuesOut(outList = outLR, parmsMat = parmsLR)

save(pLR, file = "LR_2022-02-21.RData")
```

```{r}
brks <- seq(0, 0.1, by = 0.0025)
h1 <- hist(abs(pLR$slopes), breaks = brks, freq = FALSE)
h0 <- hist(abs(pPerm$slopes), breaks = brks, freq = FALSE)
par(las = 1)
col1 <- rgb(0.8, 0, 0.8, 0.5)
col0 <- gray(0, 0.5)
plot(h1, col = col1, ylim = c(0, max(c(h0$density, h1$density))), freq = FALSE, main = "Slope values", xlab = "abs(slopes)")
plot(h0, add = TRUE, col = col0, freq = FALSE)
legend(horiz = FALSE, x ="right", pch = 15, pt.cex = 2, col = c(col0, col1), legend = c("Permutation", "Data"), box.lwd = 0)
```

```{r}
# Distribution of p values

-log(0.05/300)

cts <- c(-0.1, 0.0001, 0.001, 0.01, 0.05, 1)
hist(cut(pLR$pvals, cts, labels = FALSE), breaks = seq_along(cts) - 0.5, col = col1, ylim = c(0, 1), freq = FALSE)
hist(cut(pPerm$pvals, cts, labels = FALSE), breaks = seq_along(cts) - 0.5, add = TRUE, col = col0, freq = FALSE)

```

```{r}
# Distribution of interdecile differences
brks <- seq(0, 0.15, by = 0.005)
brks[1] <- brks[1] - 0.0001
par(las = 1)
hist(abs(pLR$tx10 - pLR$tx01), col = col1, breaks = brks, freq = FALSE, ylim = c(0, 100), main = "Interdecile difference")
hist(abs(pPerm$tx10 - pPerm$tx01), add = TRUE, col = col0, breaks = brks, freq = FALSE)
```

```{r, eval = FALSE}
for(predTransform in unique(pLR$predTransform)){
  print("-----------------------------------------------------------------------------------")
  for(varVacc in unique(pLR$varVacc)){
    for(thedate in unique(pLR$thedate)){
      for(vaccAge in unique(pLR$vaccAge)){
        for(permutation in unique(pLR$permutation)){
          sub <- pLR[pLR$vaccAge == vaccAge & pLR$thedate == thedate & pLR$predTransform == predTransform & pLR$varVacc == varVacc & pLR$permutation == permutation, ]
          print(head(sub[order(abs(sub$slopes), decreasing = TRUE), ], n = 10))
          cat("\n")
          #        print(head(sub[order(abs(sub$tx10 - sub$tx01), decreasing = TRUE), ], n = 10))
          #        cat("\n")
          plot(abs(sub$slopes), ylim = c(0, max(c(0.075, max(abs(sub$slopes))))), 
               main = paste(predTransform, thedate, vaccAge, permutation))
        }
      }
    }
  }
}

pLR$diff10.1 <- pLR$tx10 - pLR$tx01
pLR$diff10.1.se <- pLR$se10 - pLR$se01

for(predTransform in unique(pLR$predTransform)){
  print("-----------------------------------------------------------------------------------")
  for(varVacc in unique(pLR$varVacc)){
    for(thedate in unique(pLR$thedate)){
      for(vaccAge in unique(pLR$vaccAge)){
        for(permutation in unique(pLR$permutation)){
          sub <- pLR[pLR$vaccAge == vaccAge & pLR$thedate == thedate & pLR$predTransform == predTransform & pLR$varVacc == varVacc & pLR$permutation == permutation, ]
          print(head(sub[order(abs(sub$diff10.1), decreasing = TRUE), ], n = 10))
          cat("\n")
          #        print(head(sub[order(abs(sub$tx10 - sub$tx01), decreasing = TRUE), ], n = 10))
          #        cat("\n")
          yy <- abs(sub$diff10.1)
          plot(yy, ylim = c(0, 0.2), 
               main = paste(predTransform, thedate, vaccAge, permutation))
        }
      }
    }
  }
}

```

Decisions:  
-  Work with 20+ only to avoid issues with proportions of children

### Figures

```{r}
pPerm.adults <- pPerm[pPerm$vaccAge == "adults", ]
dim(pPerm.adults)

hP <- list()
brks <- seq(0, 0.2, by = 0.001)
column <- "difftx101"
for(i in 1:4){
  dt <- get(paste0("date", i))
  subP <- pPerm[pPerm$vaccAge == "adults" & pPerm$thedate == dt, ]
  hP[[i]] <- hist(abs(subP[, column]), breaks = brks)
  hP[[i]]$props <- hP[[i]]$counts / sum(hP[[i]]$counts)
  hP[[i]]$scaledcounts <- hP[[i]]$counts / max(hP[[i]]$counts)
}

i <- 1
subLR <- pLR[pLR$vaccAge == "adults" & pLR$thedate == get(paste0("date", i)), ]
plot(0, xlim = c(1, nrow(subLR)), ylim = c(0, 0.2), xaxs = "i", yaxs = "i", 
     frame.plot = FALSE, type = "n")
j <- 1:(length(brks)-1)
cols <- gray(1 - hP[[i]]$scaledcounts)
rect(xleft = rep(1, length(j)), xright = rep(nrow(subLR), length(j)), 
     ybottom = brks[j], ytop = brks[j+1], 
     border = cols, lwd = 0, col = cols)
par(xpd = TRUE)
points(abs(subLR[, column]), col = colClass[dicPred[subLR$varPred]], pch = 16)

classes <- dicPred[subLR$varPred]
aggregate(classes, FUN = length, by = list(classes))

unique(classes)

plot(0)
hP[[1]]
hP[[1]]
hist(pPerm.adults$tx10 - pPerm.adults$tx01)

```